Formalizing the Kolmogorov extension Theorem in Lean

### 2023-09-14 Manuscript/proof_outline.tex
typos

\begin{definition}[Metric and topological spaces]
  \begin{enumerate}
  \item \sloppy A pseudo-metric on $\alpha$ is a symmetric map $r : \alpha
    \times \alpha \to [0,\infty)$ satisfying the triangle inequality,
      i.e.\ $r(x,z) \leq r(x,y) + r(y,z)$ for all $x,y,z\in\alpha$.
      \\ If $r$ also satisfies $r(x,y) = 0$ iff $x=y$, we call it a
      metric.  \\ If $r : \alpha \times \alpha \to [0,\infty]$
      (i.e.\ $r(x,y) = \infty$ is allowed), we call $r$ an extended
      (pseudo-)metric.
    \item Let $r$ be an (extended pseudo-)metric. A sequence $x_1,
      x_2,... \in \alpha$ is called Cauchy if for all $\varepsilon>0$
      there is $N\in\mathbb N$ such that $r(x_n, x_m) < \varepsilon$
      for all $m,n > N$. The (extended pseudo-) metric is called
      complete if every Cauchy-sequency has a limit in
      $\alpha$.
    \item Some\footnote{We denote by $2^\alpha$ the power set of
      $\alpha$, i.e.\ the set of all subsets of $\alpha$.} $\mathcal O
      \subseteq 2^\alpha$ is called a topology, if it satisfies (i)
      $\emptyset, \alpha \in \mathcal O$, (ii) $\mathcal O$ is a
      $\pi$-system, i.e.\ it is stable under finite intersections,
      i.e.\ if $A, B \in \mathcal O$, then $A \cap B \in \mathcal O$,
      and (iii) $\mathcal O$ is stable under arbitrary unions,
      i.e.\ if $A_i \in \mathcal O$ for all $i\in\iota$ and $\iota$ is
      arbitrary, then $\bigcup_{i\in \iota} A_i \in \mathcal O$.
  \end{enumerate}
\end{definition}
\todo{reorder that next remark? I'll make a proposal}
\begin{remark}[Metric and topological spaces in {\tt mathlib}] \label{rem:uniform}\mbox{}
  \begin{enumerate}
    \item \sloppy \leanline{mathlib} provides the \leanline{class PseudoEMetricSpace (α : Type u)},
      which comes with an extended pseudo-distance
      \leanline{edist: α → α → ℝ≥0∞} (the notation \leanline{ℝ≥0∞}
      stands for $[0, \infty]$), and the properties
      \leanline{edist_self} (distance to self is zero),
      \leanline{edist_comm} (commutativity) and
      \leanline{edist_triangle} (triangle inequality).
      That class also comes with \leanline{toUniformSpace} and
      \leanline{uniformity_edist}, defining a uniform space from the
      extended pseudo-metric.
    \item A uniform space (\leanline{class UniformSpace (α : Type u)}) does not come with a
      metric, but with a filter on $\alpha \times \alpha$, which
      describes which points in $\alpha$ are near. For example, the
      diagonal of $\alpha \times \alpha$ is a subset of all sets in
      the uniformity; see \cite{james2013topologies} for details. We
      note that a uniform space with a countably generated uniformity
      filter is pseudometrizable, i.e.\ there exists a
      pseudo-metric-space structure that generates the same
      uniformity; see \leanline{UniformSpace.pseudoMetrizableSpace},
      which formalizes a result stated in
      \cite{melikhov2011metrizable}.
    \item \sloppy A topological spaces is defined in \leanline{mathlib} as \leanline{class TopologicalSpace(α : Type u)},
      which comes with a predicate \leanline{IsOpen}, and properties \leanline{isOpen_univ : IsOpen Set.univ}, \leanline{isOpen_inter} and
      \leanline{isOpen_sUnion}, which are exactly the properties of a
      topological space described above.
      An uniform spaces defines a topology, given by
      \leanline{∀ (s : Set α), IsOpen s ↔ ∀ (x : α), x ∈ s → {p | p.fst = x → p.snd ∈ s} ∈ uniformity}.
%    \item Any (extended pseudo-)metric on $\alpha$ defines a topology,
%      namely the topology generated by $\mathcal H := \{\{y:
%      r(x,y)<\varepsilon\}: x \in \alpha, \varepsilon \in
%      (0,\infty)\}$. Actually, in \leanline{mathlib}, as we have seen
%      above, every extended pseudo-metric defines a uniform space, and
%      a uniform space is a \leanline{class UniformSpace(α : Type u) extends TopologicalSpace, UniformSpace.Core} (where the
%      uniformity is defined in \leanline{UniformSpace.Core}), which
%      connects the \leanline{uniformity} with the topology using
%      \leanline{∀ (s : Set α), IsOpen s ↔ ∀ (x : α), x ∈ s → {p | p.fst = x → p.snd ∈ s} ∈ uniformity}. In particular, a
%      uniform space defines a topological space, which can be used by
%      the typeclass system.
  \end{enumerate}
\end{remark}

>>>>

\begin{definition}[Metric and topological spaces]
  \begin{enumerate}
  \item \sloppy A pseudo-metric on $\alpha$ is a symmetric map $r : \alpha
    \times \alpha \to [0,\infty)$ satisfying the triangle inequality,
      i.e.\ $r(x,z) \leq r(x,y) + r(y,z)$ for all $x,y,z\in\alpha$.
      \\ If $r$ also satisfies $r(x,y) = 0$ iff $x=y$, we call it a
      metric.  \\ If $r : \alpha \times \alpha \to [0,\infty]$
      (i.e.\ $r(x,y) = \infty$ is allowed), we call $r$ an extended
      (pseudo-)metric.
    \item Let $r$ be an (extended pseudo-)metric. A sequence $x_1,
      x_2,... \in \alpha$ is called Cauchy if for all $\varepsilon>0$
      there is $N\in\mathbb N$ such that $r(x_n, x_m) < \varepsilon$
      for all $m,n > N$. The (extended pseudo-) metric is called
      complete if every Cauchy-sequency has a limit in
      $\alpha$.
    \item Some\footnote{We denote by $2^\alpha$ the power set of
      $\alpha$, i.e.\ the set of all subsets of $\alpha$.} $\mathcal O
      \subseteq 2^\alpha$ is called a topology, if it satisfies (i)
      $\emptyset, \alpha \in \mathcal O$, (ii) $\mathcal O$ is a
      $\pi$-system, i.e.\ it is stable under finite intersections,
      i.e.\ if $A, B \in \mathcal O$, then $A \cap B \in \mathcal O$,
      and (iii) $\mathcal O$ is stable under arbitrary unions,
      i.e.\ if $A_i \in \mathcal O$ for all $i\in\iota$ and $\iota$ is
      arbitrary, then $\bigcup_{i\in \iota} A_i \in \mathcal O$.
  \end{enumerate}
\end{definition}

\begin{remark}[Metric and topological spaces in {\tt mathlib}] \label{rem:uniform}\mbox{}
  \begin{enumerate}
    \item \sloppy \leanline{mathlib} provides the \leanline{class PseudoEMetricSpace (α : Type u)},
      which comes with an extended pseudo-distance
      \leanline{edist: α → α → ℝ≥0∞} (the notation \leanline{ℝ≥0∞}
      stands for $[0, \infty]$), and the properties
      \leanline{edist_self} (distance to self is zero),
      \leanline{edist_comm} (commutativity) and
      \leanline{edist_triangle} (triangle inequality).
      That class also comes with \leanline{toUniformSpace} and
      \leanline{uniformity_edist}, defining a uniform space from the
      extended pseudo-metric.
    \item A uniform space (\leanline{class UniformSpace (α : Type u)}) does not come with a
      metric, but with a filter on $\alpha \times \alpha$, which
      describes which points in $\alpha$ are near. For example, the
      diagonal of $\alpha \times \alpha$ is a subset of all sets in
      the uniformity; see \cite{james2013topologies} for details. We
      note that a uniform space with a countably generated uniformity
      filter is pseudometrizable, i.e.\ there exists a
      pseudo-metric-space structure that generates the same
      uniformity; see \leanline{UniformSpace.pseudoMetrizableSpace},
      which formalizes a result stated in
      \cite{melikhov2011metrizable}.
    \item \sloppy A topological space is defined in \leanline{mathlib} as \leanline{class TopologicalSpace(α : Type u)},
      which comes with a predicate \leanline{IsOpen}, and properties \leanline{isOpen_univ : IsOpen Set.univ}, \leanline{isOpen_inter} and
      \leanline{isOpen_sUnion}, which are exactly the properties of a
      topological space described above.
      A uniform space defines a topology, given by
      \leanline{∀ (s : Set α), IsOpen s ↔ ∀ (x : α), x ∈ s → {p | p.fst = x → p.snd ∈ s} ∈ uniformity}.
%    \item Any (extended pseudo-)metric on $\alpha$ defines a topology,
%      namely the topology generated by $\mathcal H := \{\{y:
%      r(x,y)<\varepsilon\}: x \in \alpha, \varepsilon \in
%      (0,\infty)\}$. Actually, in \leanline{mathlib}, as we have seen
%      above, every extended pseudo-metric defines a uniform space, and
%      a uniform space is a \leanline{class UniformSpace(α : Type u) extends TopologicalSpace, UniformSpace.Core} (where the
%      uniformity is defined in \leanline{UniformSpace.Core}), which
%      connects the \leanline{uniformity} with the topology using
%      \leanline{∀ (s : Set α), IsOpen s ↔ ∀ (x : α), x ∈ s → {p | p.fst = x → p.snd ∈ s} ∈ uniformity}. In particular, a
%      uniform space defines a topological space, which can be used by
%      the typeclass system.
  \end{enumerate}
\end{remark}
